---
title: "SDS/MTH 220: Intro to Probability and Statistics"
author: "Sec01: Albert Y. Kim & David Rockoff. Sec02: Katherine Kinnaird & Will Hopper"
date: "Last updated on `r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
    df_print: paged
    includes:
      in_header: "favicon.html"
---

<style>
h1{font-weight: 400;}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo=TRUE, 
  message=FALSE, 
  warning=FALSE, 
  eval=FALSE, 
  fig.width=16/2, 
  fig.height=9/2
)
library(tidyverse)
library(lubridate)
library(stringr)
library(forcats)
library(moderndive)

# Formating packages:
library(knitr)
library(kableExtra)
library(scales)

# Data packages:
library(fivethirtyeight)
library(nycflights13)
library(gapminder)
library(babynames)

# Etc
library(here)
library(patchwork)

# Set seed value of random number generator to get "replicable" random numbers.
set.seed(76)

# Example points for boxplots
example <- tibble(
  values = c(1, 3, 5, 6, 7, 8, 9, 12, 13, 14, 15, 30)
)

# fruit basket example for data wrangling
fruit_basket <- tibble(
  type = c("mango", "kiwi", "mango", "grape", "grape", "mango"),
  `price in $` = c(2, 3, 1, 3, 1, 3)
)
```



***



# Schedule 

<iframe src="https://docs.google.com/spreadsheets/d/e/2PACX-1vQM-5FVt5BKJZsfElEV19NwD2MXGgSVUfGLB7dUoD0MbCU6Dkzqx66UMxbgqMaiS6H8CP8numV5Xaf-/pubhtml?gid=346882447&amp;single=true&amp;widget=true&amp;headers=false" width="100%" height="1015"></iframe>

<!--
{target="_blank"}
-->



***



# Lec 09: Fri 2/14

## Announcements

## Recap

## Chalk talk

<!--
Write out table of points & regression line values for data in Fig 5.13
-->

* Connection between `estimate` in regression table and plot of regression.
* Interpretation of:
    + Fitted intercept $b_0$: mathematical versus practical interpretations.
    + Fitted slope $b_1$: language matters!
* Obtaining a regression table takes two steps:
    + Create/fit the model using `lm()` and formula interface `y ~ x`
    + Ask for the table using `get_regression_table()`



***



# Lec 08: Wed 2/12

## Announcements

* We move to new chapter in ModernDive: data modeling

## Recap

## Chalk talk

1. 5.4.1: Guessing the correlation game
1. Two types of variables: outcome/response $y$ and explanatory/predictor variables $x$
1. Example models:
    + Temp in F vs temp in C
    + Coffee vs sleep
1. Correlation coefficients: Quantify the strength of linear assocation. 



***



# Lec 07: Mon 2/10

## Announcements

* Today's lecture is last of pink topics in calendar (data science). We start blue topics (regression) on Wednesday.

## Recap

## Chalk talk

* Using same fruit basket example from chalk talk in Lec 05:  
```{r, echo=FALSE, eval=TRUE}
fruit_basket
```
    1. `mutate()` new variables from existing variables: price in cents
    1. Reorder i.e. `arrange()` rows of a data frame in alpha-numerical order of a variable: in order of price. Default is ascending order, need to specify `desc()`
    1. `select()` variables/columns. Analogous to `filter()` rows.



***


# Lec 06: Fri 2/7

## Announcements

<!--
* Bert only: Emphasize problem set policy. "Turn in what you have."
* Bert only: Google doc exercise: https://docs.google.com/document/d/14PBmTU_wBTEI8j1lqZWFqFI3ETAD6N5qao-2ADQRTTM/edit

Write down 3 things you did for this class this week
Turn to a neighbor, come up with a list of 5 things
Then we did the google doc
-->

## Recap

<!--
* Bert only based on questions in his section: Distinguishing between
    + English use of "or" and "and" versus
    + Computer programming use of "or" and "and"
-->

## Chalk talk

* Using same fruit basket example from chalk talk in Lec 05:  
```{r, echo=FALSE, eval=TRUE}
fruit_basket
```
    1. "Summary functions" are many-to-one functions to compute "summary statistics," like `mean()` & `median()`
    1. Using summary functions to then `summarize()` rows
    1. Setting "group meta-data" of a data frame using `group_by()`, then `summarize()` rows. In other words, `group_by()` by itself does not change the "data", rather only the "meta-data"
* Normal distribution: One summary statistic of spread, standard deviation `sd()`, has a particular interpretation when distribution of a numerical variable follows a normal AKA bell curve



***




# Lec 05: Wed 2/5

## Announcements

* Posted single calendar of all instructor and lab instructor office hours on Moodle.
* Posted [badge challenges](badge_challenges.html)
* Managing Slack notifications:
    + We suggest but don't require installing Desktop App. 
    + If not, at the very least, keep email notifications on.
* How to approach in-class readings.
    
<!--
* Bert only: "not equals" and "bang equals"
-->

## Recap 

First, let's compare the boxplot you created in Lec04 with the original points. See how they align.

```{r, echo=FALSE, eval=TRUE}
plot1 <- ggplot(example, aes(x=factor(1), y=values)) +
  geom_point() +
  labs(title = "Points") +
  theme(
    axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()
  )

plot2 <- ggplot(example, aes(x=factor(1), y=values)) +
  geom_boxplot() +
  labs(title = "Boxplot") +
  theme(
    axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()
  )
plot1 + plot2
```

Second, let's compare two visualizations of the distribution of the numerical variable temperature *split by* month:

* Faceted histogram in ModernDive [Figure 2.13](https://moderndive-bert.netlify.com/2-viz.html#facets){target="_blank"}
* Side-by-side boxplot in ModernDive [Figure 2.18](https://moderndive-bert.netlify.com/2-viz.html#boxplots){target="_blank"}

```{r, echo=FALSE, eval=TRUE}
ggplot(data = weather, mapping = aes(x = temp)) +
  geom_histogram(binwidth = 5, color = "white") +
  facet_wrap(~ month)
ggplot(data = weather, mapping = aes(x = factor(month), y = temp)) +
  geom_boxplot()
```

Which do you prefer?

Third, the idea of "assigning AKA saving something in R" using the arrow `<-`. Ex: `x <- 5`. You do this so you can access it again later, either to

1. View its contents: Type `x` in the console or RMarkdown.
1. Modify it's contents: Type `x <- x + 1` or RMarkdown. What is now saved in `x`?


## Chalk talk

We now start data wrangling! For our chalk talks, we'll use the following "fruit basket" example:

```{r, echo=FALSE, eval=TRUE}
fruit_basket
```

<br/>
Topics:

1. Define the "pipe operator" `%>%`, pronounced *then*
1. Logical AKA boolean operators in computer programming
1. `filter()` rows that match a criteria



***



# Lec 04: Mon 2/3

## Announcements

* 

## Chalk talk

Boxplots! Powerful, but tricky!

Say we want to study the distribution of the following `r nrow(example)` values which are pre-sorted: 

> `r example %>% pull(values)`

They have the following *summary statistics*. A summary statistic is a single numerical value *summarizing* many values. Examples include the immediately obvious mean AKA average and median. Other less immediately obvious examples include:

* Quartiles (1st, 2nd, and 3rd) that cut up the data into 4 parts, each containing roughly one quarter = 25% of the data
* Minimum & maximum
* Interquartile-range (IQR): the distance between the 3rd and 1st quartiles

Min. | 1st Quartile | Median = 2nd Quartile  | 3rd Quartile  |  Max. | IQR
---- | ------- | ------  |  ------- | --- | ---
`r min(example$values)`  |  `r quantile(example$values, probs=0.25, type = 2)` |  `r median(example$values)`  |  `r quantile(example$values, probs=0.75, type=2)`  | `r max(example$values)` | 8 = 13.5 - 5.5


Steps to constructing a boxplot:

1. Draw box
1. Throw out whiskers 1.5 x IQR from either end of box
1. Mark outliers
1. Pull in the ends of both whiskers towards the box until an observed value is hit.



***



# Lec 03: Fri 1/31

## Announcements


## Chalk talk

1. Draw histogram by hand, emphasizing that bins correspond to intervals (right-edge inclusive).
1. Adjust the binning structure of the previous hand-drawn histogram two ways:
    + Adjusting the binwidth
    + Adjusting the number of bins
1. Facets split one graphic by the values of another variable (that does not have too many unique values).



***



# Lec 02: Wed 1/29

## Announcements

1. Fatima Keita is interested in organizing a group for students of color within SDS, called SDS coalition of color (SDSCC). She has also put together a survey to gather students of color's experiences within SDS courses and the program more generally. If you identify as a student of color and you've taken at least one SDS course, please take a moment to complete Fatima's [SDSCC Preliminary Interest Survey](https://www.surveymonkey.com/r/2YZ785Y){target="_blank"}. The idea is that these experiences would then be anonymously shared with the SDS faculty in hopes that we can work to improve inclusivity within the program and curriculum. If you are interested in joining, please email Fatima directly.
1. Show example `#questions` on Slack.
1. In-class demo of Gradescope.
1. This week's problem set has three elements
    + Submit PDF on Gradescope with questions tagged
    + Fill out Intro Survey Google Form
    + Submit on Moodle reflection piece which builds off in-class reflection exercise.
    
    
## Chalk Talk

1. Draw scatterplot by hand
1. Define of Grammar of Graphics
1. Write example `ggplot()` code
1. Explain what jittering does to graphic and what it does not do to original data



***



# Lec 1: Mon 1/27

## Announcements

* Introductions
* If you're having registration issues:
    + Please keep in mind that SDS201 is a course that also satisfies the intro stats pre-requisite for many majors, including PSY, GOV, BIO, NSC, ENV, and SDS. Please consider that option as it currently has lots of room.
    + If you are not registered for this course, fill out the waitlist [Google Form](https://docs.google.com/forms/d/e/1FAIpQLSdDfcVmrz8p2rmS2BXqT6tv3a-Syzn2gmIG88h4CYQzRvKltw/viewform){target="_blank"} before Thursday 5pm.
    + We will start informing students and signing green slips on Friday afternoon.
    + If you have previously put your name on the above waitlist Google Form but no longer need a spot, we would very much appreciate a message letting us know. This will help us and your peers a lot.
* Syllabus 
    + Honor code discussions
    + Reflection exercise
    
    
## Lecture

Here is the typical lecture flow:

* Start with a 10-15 minute "chalk talk". We do this to motivate the day's topics and warn you about potential pitfalls in understanding
* Followed by in-class ModernDive readings listed in the above schedule. We do this so you can read at your own pace, talk to your peers, and ask us questions. While you do not need to submit your "Learning Check" answers, we *highly* recommend you write them down and verify your answers in Appendix D of ModernDive.

Today's topics:

* Introduction to Slack: A brief demo
* Introduction to R, RStudio, and R packages: ModernDive readings
