---
title: "SDS/MTH 220: Intro to Probability and Statistics"
author: "Sec01: Albert Y. Kim & David Rockoff. Sec02: Katherine Kinnaird & Will Hopper"
date: "Last updated on `r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
    df_print: kable
    includes:
      in_header: "favicon.html"
---

<style>
h1{font-weight: 400;}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo=TRUE, 
  message=FALSE, 
  warning=FALSE, 
  eval=FALSE, 
  fig.width=16/2, 
  fig.height=9/2
)
library(tidyverse)
library(stringr)
library(knitr)
library(lubridate)
library(forcats)
library(fivethirtyeight)
library(nycflights13)
library(gapminder)
library(babynames)
library(scales)
library(moderndive)
library(here)
library(kableExtra)
# devtools::install_github("hadley/emo")
library(emo)
library(patchwork)

# Set seed value of random number generator to get "replicable" random numbers.
set.seed(76)
```



***



# Schedule 

<iframe src="https://docs.google.com/spreadsheets/d/e/2PACX-1vQM-5FVt5BKJZsfElEV19NwD2MXGgSVUfGLB7dUoD0MbCU6Dkzqx66UMxbgqMaiS6H8CP8numV5Xaf-/pubhtml?gid=346882447&amp;single=true&amp;widget=true&amp;headers=false" width="100%" height="1015"></iframe>

<!--
{target="_blank"}
-->



***



# Lec 05: Wed 2/5

## Annoucements

* Recap: 
    + points vs boxplots
    + comparing boxplots to faceted histograms
    + The idea of "saving something to a variable". You do this to access it again later. You view its contents by just typing variable name in console

## Chalk talk

* Define `fruit_basket`
* pipe
* `filter()`




# Lec 04: Mon 2/3

## Annoucements


## Chalk talk

Boxplots! Powerful, but tricky!

```{r, eval=TRUE, echo=FALSE}
library(ggplot2)
library(dplyr)
example <- tibble(
  values = c(1, 3, 5, 6, 7, 8, 9, 12, 13, 14, 15, 30)
)
```

Say we want to study the distribution of the following `r nrow(example)` values which are pre-sorted: 

> `r example %>% pull(values)`

They have the following *summary statistics*. A summary statistic is a single numerical value *summarizing* many values. Examples include the immediately obvious mean AKA average and median. Other less immediately obvious examples include:

* Quartiles (1st, 2nd, and 3rd) that cut up the data into 4 parts, each containing roughly one quarter = 25% of the data
* Minimum & maximum
* Interquartile-range (IQR): the distance between the 3rd and 1st quartiles

Min. | 1st Quartile | Median = 2nd Quartile  | 3rd Quartile  |  Max. | IQR
---- | ------- | ------  |  ------- | --- | ---
`r min(example$values)`  |  `r quantile(example$values, probs=0.25, type = 2)` |  `r median(example$values)`  |  `r quantile(example$values, probs=0.75, type=2)`  | `r max(example$values)` | 8 = 13.5 - 5.5


Steps to constructing a boxplot:

1. Draw box
1. Throw out whiskers 1.5 x IQR from either end of box
1. Mark outliers
1. Pull in the ends of both whiskers towards the box until an observed value is hit.




<!--
We'll reveal this and the output below on Lec5 on Wed:

Let's compare the points and the corresponding boxplot side-by-side with the values on the $y$-axis matching:
-->
```{r, echo=FALSE, eval=FALSE}
plot1 <- ggplot(example, aes(x=factor(1), y=values)) +
  geom_point() +
  labs(title = "Points") +
  theme(
    axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()
  )

plot2 <- ggplot(example, aes(x=factor(1), y=values)) +
  geom_boxplot() +
  labs(title = "Boxplot") +
  theme(
    axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()
  )
plot1 + plot2
```




***



# Lec 03: Fri 1/31

## Annoucements


## Chalk talk

1. Draw histogram by hand, emphasizing that bins correspond to intervals (right-edge inclusive).
1. Adjust the binning structure of the previous hand-drawn histogram two ways:
    + Adjusting the binwidth
    + Adjusting the number of bins
1. Facets split one graphic by the values of another variable (that does not have too many unique values).



***



# Lec 02: Wed 1/29

## Annoucements

1. Fatima Keita is interested in organizing a group for students of color within SDS, called SDS coalition of color (SDSCC). She has also put together a survey to gather students of color's experiences within SDS courses and the program more generally. If you identify as a student of color and you've taken at least one SDS course, please take a moment to complete Fatima's [SDSCC Preliminary Interest Survey](https://www.surveymonkey.com/r/2YZ785Y){target="_blank"}. The idea is that these experiences would then be anonymously shared with the SDS faculty in hopes that we can work to improve inclusivity within the program and curriculum. If you are interested in joining, please email Fatima directly.
1. Show example `#questions` on Slack.
1. In-class demo of Gradescope.
1. This week's problem set has three elements
    + Submit PDF on Gradescope with questions tagged
    + Fill out Intro Survey Google Form
    + Submit on Moodle reflection piece which builds off in-class reflection exercise.
    
    
## Chalk Talk

1. Draw scatterplot by hand
1. Define of Grammar of Graphics
1. Write example `ggplot()` code
1. Explain what jittering does to graphic and what it does not do to original data



***



# Lec 1: Mon 1/27

## Announcements

* Introductions
* If you're having registration issues:
    + Please keep in mind that SDS201 is a course that also satisfies the intro stats pre-requisite for many majors, including PSY, GOV, BIO, NSC, ENV, and SDS. Please consider that option as it currently has lots of room.
    + If you are not registered for this course, fill out the waitlist [Google Form](https://docs.google.com/forms/d/e/1FAIpQLSdDfcVmrz8p2rmS2BXqT6tv3a-Syzn2gmIG88h4CYQzRvKltw/viewform){target="_blank"} before Thursday 5pm.
    + We will start informing students and signing green slips on Friday afternoon.
    + If you have previously put your name on the above waitlist Google Form but no longer need a spot, we would very much appreciate a message letting us know. This will help us and your peers a lot.
* Syllabus 
    + Honor code discussions
    + Reflection exercise
    
    
## Lecture

Here is the typical lecture flow:

* Start with a 10-15 minute "chalk talk". We do this to motivate the day's topics and warn you about potential pitfalls in understanding
* Followed by in-class ModernDive readings listed in the above schedule. We do this so you can read at your own pace, talk to your peers, and ask us questions. While you do not need to submit your "Learning Check" answers, we *highly* recommend you write them down and verify your answers in Appendix D of ModernDive.

Today's topics:

* Introduction to Slack: A brief demo
* Introduction to R, RStudio, and R packages: ModernDive readings
