---
title: "Getting to know RStudio"
author: "Cosme Fulanito"
date: "1/25/2022"
output:
  pdf_document: default
  html_document: default
---

## This is a header
Text goes here in the white part

## This is another header
Below is a code chunk! Code goes in here:

```{r}
x <- 6
```

```{r}
favorite_movie <- "Star_Wars"
```

```{r}
v <- c(2, 4, 6)
v*3
```

## Exercise 1

```{r}
y <- 7
z <- x*y
```

## Exercise 2

```{r}
6+3
```

The answer appears in the console

## Exercise 3

```{r}
a <- 6 + 3
```

The answer does not appear.

The object a shows up in the Environment pane.

```{r}
a
```

The answer now appears in the console.


## Exercise 4

```{r}
a^2
```

It raises the value of a to the power of two i.e. squares it. 


## Exercise 5

```{r}
sum(a, x, y)
```

The sum function adds the values of numbers together



## Exercise 6

All the objects disappeared! They got deleted!



## Exercise 7

All the objects reappeared! We were able to recreate them!



## Exercise 8

```{r}
v + 2
```

This code adds 2 to all the original values of v


## Exercise 9

```{r}
music <- c("bluegrass", "funk", "folk")
```

In the Environment pane, second column, we see chr which is short for character. Also all the values of music are surrounded by quotation marks


## Exercise 10

*Italicize like this*

**Bold like this**

A superscript: R^2^

## Exercise 11

All the versions of R are named after Charlie Brown comics! For example, R version 4.0.3 released on 2020-10-10 is named "Bunny-Wunnies Freak Out" https://peanuts.fandom.com/wiki/Bunny_Wunny





